Supplemental Materials for ``Universals of word order reflect optimization of grammars for efficient communication''

### 2019-11-12 writeup/3-paper-names.tex
code

\begin{table}[ht]
	\begin{center}
    \includegraphics[width=0.45\textwidth]{table1.pdf}
	\end{center}
	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples: Across the world, the orders of different constituents are strikingly correlated with that of verb and object.
	Selection is based on a more recent typological study by Dryer~\cite{dryer1992greenbergian}, restricted to those correlations that are annotated in available corpus data. See SI Appendix, Section S1 for more on Greenberg correlations.
	}\label{table:corr-dryer}
\end{table}

>>>>

\begin{table}[ht]
	\begin{center}
\begin{tabular}{|c|ll|ll|}
	\hline
	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Turkish, ...)} \\ \hline\hline
	& \multicolumn{2}{c|}{Correlates with...} & \multicolumn{2}{c|}{Correlates with...}  \\
	&	Verb & Object     & Object & Verb    \\ \hline\hline
	& kataba & ris{\= a}la 	&tegami-o & kaita \\
	&	\emph{wrote} & \emph{letter} & 	\emph{letter} & \emph{wrote} \\ \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    {\d s}ad{\= i}q       &	tomodachi & ni \\ % G3, G4
	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&k{\= a}na    &    {\d s}ad{\= i}q         &	tomodachi & datta \\ % not in Greenberg
	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak- & -udesho \\ % G16, G13 TODO this should only be inflected auxiliaries, replace Arabic
	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& {\d s}ad{\= i}q  &    John    & 	John no & tomodachi \\ % G2, G23
	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uh{\= a}       & 	anata-ga yonda & hon \\ % G24
	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tu{\d s}il        & 	toochaku suru & koto \\ % 
	&	\emph{that} & \emph{she arrives}  &	\emph{arrives} & \emph{that} \\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'il{\= a} lmadrasa     & 	gakkoo ni & itta \\ % G22
	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'ur{\= i}d    &    'an 'ugh{\= a}dir         & 	ik- & -itai \\ % G15
	& \emph{wants}   &  \emph{to leave}  &	\emph{to go} & \emph{want} \\ \hline 
\end{tabular}
	\end{center}
	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples: Across the world, the orders of different constituents are strikingly correlated with that of verb and object.
	Selection is based on a more recent typological study by Dryer~\cite{dryer1992greenbergian}, restricted to those correlations that are annotated in available corpus data. See SI Appendix, Section S1 for more on Greenberg correlations.
	}\label{table:corr-dryer}
\end{table}

### 2019-11-12 writeup/pnas-word-order-si.tex
code

\item Correlation {\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}} is formalized using \emph{obl}, which covers not only PPs and standards of comparison, but also adjunct NPs.
In the WALS evaluation (Table~\ref{tab:grammars-wals}), the dominant order of \emph{nmod} agrees with that annotated for obliques in all 18 languages for which WALS has an entry.

@@

We obtained the posterior density plots in Table 2 (Main Paper) and in Figure (\ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
As the logistic transformation is inverse to the logit transform (\ref{eq:mixed-effects}), this corresponds to the posterior distribution of the prevalence (between 0.0 and 1.0) of each correlation, controlling for languages and language families.

@@

\begin{table*} 
	\begin{center}	
\includegraphics[width=0.8\textwidth]{si-table-perrel-1a-1.png}
\end{center}
\caption{Predictions on UD relations with predictions from the typological literature.
The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
The `Real' column provides the prevalence among the 51 languages in the Universal Dependencies data.
We provide posterior prevalences for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone, obtained from the Bayesian mixed-effects analysis controlling for languages and language families (as in Table 2 of the main paper).
In the last column, we indicate what prevalence is expected according to the typological literature. 
}
\label{tab:all-predictions-1}
\end{table*}

>>>>

\item Correlation {\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}} is formalized using \emph{obl}, which covers not only PPs and standards of comparison, but also adjunct NPs.
In the WALS evaluation (Table~\ref{tab:grammars-wals}), the dominant order of \emph{obl} agrees with that annotated for obliques in all 18 languages for which WALS has an entry.

@@

We obtained the posterior density plots in Figure 6 (Main Paper) and in Figure (\ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
As the logistic transformation is inverse to the logit transform (\ref{eq:mixed-effects}), this corresponds to the posterior distribution of the prevalence (between 0.0 and 1.0) of each correlation, controlling for languages and language families.

@@

\begin{table*} 
	\begin{center}	
\includegraphics[width=0.8\textwidth]{si-table-perrel-1a-1.png}
\end{center}
\caption{Predictions on UD relations with predictions from the typological literature.
The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
The `Real' column provides the prevalence among the 51 languages in the Universal Dependencies data.
We provide posterior prevalences for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone, obtained from the Bayesian mixed-effects analysis controlling for languages and language families (as in Figure 6 of the main paper).
In the last column, we indicate what prevalence is expected according to the typological literature. 
}
\label{tab:all-predictions-1}
\end{table*}
